### 2023-07-18 article/code/SVGFunc.py.tex

# Function that changes the energy bar height
# "soup" is the SVG data
# "element_id" is the id of the energy bar that will be resized
# "new_height" is the new height of the energy bar
def _changeElementHeight(self, soup, element_id, new_height):

 # Find the element
 element = soup.find(id=element_id)

 # If it exists updates its height and paint it of 
 # black because initially the bars are transparent
 if element is not None:
   element['height'] = str(new_height/1000)
   element["style"] = 'fill:#000000'

# Function that paint the electrodes in the SVG image
# "soup" is the SVG data
# "sensorsTags" is the tag of the electrode to be painted
def _paintSensor(self, soup, sensorsTags):

 # Find the element
 element = soup.find(id=sensorsTags)

 # If it exists updates its color
 if element is not None:
   color = self.ddata.sensorToColorRelationship[sensorsTags]
   element["style"] = f'fill:{color}'