### 2023-07-18 article/code/dataProcessing.py.tex

# Loads the EEG section into a dataframe
self.dataframe = pd.DataFrame(mat['EEG'])

# Joins the columns into its respectives electrodes
self._joinIntoArray(0, 4096, 'F3', self.dataframe)
self._joinIntoArray(0, 4096, 'F4', self.dataframe)
self._joinIntoArray(0, 4096, 'C3', self.dataframe)
self._joinIntoArray(0, 4096, 'C4', self.dataframe)
self._joinIntoArray(0, 4096, 'P3', self.dataframe)
self._joinIntoArray(0, 4096, 'P4', self.dataframe)

# Joins the columns into its respectives classes
self._joinIntoValue(0,1,'Modality',self.dataframe)
self._joinIntoValue(0,1,'Estimuli',self.dataframe)
self._joinIntoValue(0,1,'Artifact',self.dataframe)